Multiple testing with logical dependencies

### 2013-08-13 paper/draft.tex
forgot to push no idea?
put the OM result into theorem form

Now observe that each hypothesis $H_i$ corresponds to a hyperplane
$h_i = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} = b_i\}$ that divides
$\mathbb{R}^m$ into two regions $h_i^+ = \{\bs{x} \in \mathbb{R}^m:
\bs{A}_i^t\bs{x} < b_i\}$ and $h_i^- = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} >
b_i\}$ giving rise to a sign function:
\begin{equation}
  \label{eq:sign}
  \delta_i(\bs{x}) = \left\{
\begin{array}{ll}
  + & \text{if } \bs{x} \in h_i^+\\
  0 & \text{if } \bs{x} \in h_i \\
  - & \text{if } \bs{x} \in h_i^-\\
\end{array}\right.
\end{equation}

Further for every $\bs{x} \in \mathbb{R}^m$ we have  sign vector
$\bs{\delta}{\bs{x}}= (\delta_1(\bs{x}),...,\delta_n(\bs{x}))$. Now
for each $i$ and some parameter value $\bs{\theta} \in \bs{\Theta} = \R^m$,
$\delta_i(\bs{\theta})$ tells us whether the corresponding value is under the null
hypotheses on its border, or under the alternative. Accordingly,
$\bs{\delta}(\bs{\theta})$ tells us in which combination of true and
false null hypotheses $H_i$ correspond to $\bs{\theta}$. Let
\begin{equation}
\label{eq:faces}
\mathcal{F}(\bs{A},\bs{b} = \{\bs{\delta}(\bs{x}): \bs{x} \in \R^m\}
\end{equation}

and for some $F \in \mathcal{F}$ define
\begin{equation}
  \label{eq:partition}
  \Theta_F = \{\bs{x} \in \R^m: \bs{\delta}(\bs{x}) = F\}
\end{equation}

then we have $\bigcup_{F \in \mathcal{F}} \Theta_F = \bs{\Theta} =
\R^m$ and for any $\Theta_F \cap \Theta_G = \emptyset$ for any $F \neq
G$. Consequently, $\{\Theta_F:F \in \mathcal{F}$ represents a
partition of the parameter space and each element in $\mathcal{F}$
uniquely identifies some {\em non-empty} element of that partition. Consequently,
following the Partitioning Principle \cite[]{finner2002partitioning}
if we have level $\alpha$ tests $\varphi_F(X)$ for each $F \in
\mathcal{F}$ a test that rejects $H_i$ for observed values $X$ if
$\varphi_F(X) = 1$ for all $F$ with $F_i \geq 1$ controls the FWER in
the strong sense. For example one may use  weighted Bonferroni tests
\begin{equation}
  \label{eq:bonferroni}
  \varphi_F = \indf{\min_{i:F_i \geq 0}p_i \leq w_{F,i}}
\end{equation}
where the $w_{F,i}$ are can be computed using a graphical approach.

@@

\begin{definition}
  Consider a family of hypotheses $\mathscr{H} =$
\end{definition}

>>>>

\subsection{Arrangements of hyperplanes}
\label{sec:hyperplanes}

Now observe that each hypothesis $H_i$ corresponds to a hyperplane
$h_i = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} = b_i\}$. The
family of null hypotheses $\mathscr{H}$ therefore is full determined
by an arrangement of hyperplanes $\mathscr{A} = \{h_i:i=
\{1,...,n\}\}$.

\begin{definition}
  An arrangement of hyperplanes is said to be regular if there is a
  subset of $\mathscr{A}$ whose intersection is a single point
\end{definition}

Two important cases for non-regular hyperplane arrangements come to
mind. First, all pairwise intersections of hyperplanes are empty. If
that is the case then all hypotheses are nested. Such a case is very
simple an appropriate testing algorithm would be a hierarchical
approach starting with the {\em smallest} null hypotheses. Second, all
hypotheses intersect in more than one point. In this case, however, we
know from basic linear algebra that the intersection is an affine
subspace. Consequently one can find a hyperplane that is orthogonal to
this subspace onto which we can project $\mathscr{A}$ to get a central
regular arrangement of lower dimensionality. An example of this would
be the family of hypotheses defined by contrast matrix

\begin{equation}
  \label{eq:example:ordered}
  \bs{A} = \left\{
    \begin{array}{ccc}
      1 & -1 & 0 \\
      1 & 0 & -1 \\
      0 & 1 & -1 
    \end{array}
  \right\}
\end{equation}

and vector $\bs{b} = (0,0,0)^t$. Here all points of the form $(x,x,x)$, $x
\in \R^3$ are a solution to the corresponding system of linear
equations. On can now project use the plane $x_1 + x_2 + x_3 = 0$ as a
ground and represent the hyperplanes corresponding to $\bs{A}$, and
$\bs{b}$ by the lines $l_1,l_2,l_3$
{\tt Derive the appropriate projection onto the plane rectangular to
  $(1,1,1)x=0$ and put here}

Notice that each hyperplane $h_i$ divides
$\mathbb{R}^m$ into two regions $h_i^+ = \{\bs{x} \in \mathbb{R}^m:
\bs{A}_i^t\bs{x} < b_i\}$ and $h_i^- = \{\bs{x} \in \mathbb{R}^m: \bs{A}_i^t\bs{x} >
b_i\}$. For any point $x$ in the parameter space and elementary
hypotheses $H_i$ we define sign function: 
\begin{equation}
  \label{eq:sign}
  \delta_i(\bs{x}) = \left\{
\begin{array}{ll}
  + & \text{if } \bs{x} \in h_i^+\\
  0 & \text{if } \bs{x} \in h_i \\
  - & \text{if } \bs{x} \in h_i^-\\
\end{array}\right.
\end{equation}

{\em I.e.,} $\delta_i$ is takes value $-$ if $H_i$ at $x$ is false and
$0$ or $+$ otherwise.

Now, for every $\bs{x} \in \mathbb{R}^m$ we define a sign vector
$\bs{\delta}{\bs{x}}=
(\delta_1(\bs{x}),...,\delta_n(\bs{x}))$. Accordingly,
$\bs{\delta}(\bs{x})$ tells us to which 
combination of true and false null hypotheses some $\bs{x}$
belongs. Let
\begin{equation}
\label{eq:faces}
\mathcal{F}(\mathscr{H}) = \{\bs{\delta}(\bs{x}): \bs{x} \in \R^m\}
\end{equation}

denote the set of all sign patterns defined by 
$\mathscr{H}$.  Further it
provides a multiple testing procedure for $\mathscr{H}$ that controls
the FWER in the strong sense:

\begin{theorem}
  Consider a family of general contrast hypotheses $\mathscr{H} =
  \{H_i: \bs{A_i}^t\bs{\mu} \leq b_i,i \in \{1,...,n\}\}$ and let
  $\mathscr{F}(\mathscr{H}) = \{ \bs{delta}(\bs{x}): \bs{x} \in
  \R^m\}$ denote the corresponding set of sign patterns. Consider a
  set of decision rules $\varphi_F: \R^n \rightarrow \{0,1\}$ for
  which hold $P(\varphi_F = 1 | H_i: F_i \geq 0) \leq \alpha$. Then a
  corresponding decision rule for each elementary hypotheses $H_i$ is
  given by 
  \begin{equation}
    \label{eq:signtest}
    \psi_i = \min_{F: F_i \geq 0} \varphi_F,
  \end{equation}
  which controls the FWER at level $\alpha$ in the strong sense. 
\end{theorem}

\begin{proof}
  For all $F \in \mathcal{F}$ define
  \begin{equation}
    \label{eq:partition}
    \Theta_F = \{\bs{x} \in \R^m: \bs{\delta}(\bs{x}) = F\},
  \end{equation}

  which is a subset of the parameter space. We have $\bigcup_{F \in
    \mathcal{F}} \Theta_F = \bs{\Theta} = \R^m$ and for any $\Theta_F
  \cap \Theta_G = \emptyset$ for any $F \neq G$. Consequently,
  $\{\Theta_F:F \in \mathcal{F}$ represents a partition of the
  parameter space and each element in $\mathcal{F}$ uniquely
  identifies some {\em non-empty} element of that
  partition. Consequently, following the Partitioning Principle
  \cite[]{finner2002partitioning} if we have level $\alpha$ tests
  $\varphi_F(X)$ for each $F \in \mathcal{F}$ a test that rejects
  $H_i$ for observed values $X$ if $\varphi_F(X) = 1$ for all $F$ with
  $F_i \geq 1$ controls the FWER in the strong sense.
\end{proof}

For example one may use  weighted Bonferroni tests
\begin{equation}
  \label{eq:bonferroni}
  \varphi_F = \indf{\min_{i:F_i \geq 0}p_i \leq w_{F,i}}
\end{equation}
where the $w_{F,i}$ are can be computed using a graphical approach.

@@
